### 2020-05-08 Clustering/2019_qb_preparation.Rmd
initial commit of files

---
title: "K-Means Clustering of 2019 NFL Quarterbacks, Running Backs and Receivers"
author: "Philip Ipe"
date: "4/3/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(lubridate)
library(tidyverse)
library(knitr)
```

## Data Scraping and Preparation

Scraping Advanced Passing Data from 2019 from Pro-Football Reference
```{r scraping_passing}
player_adv_stats_url <- "https://www.pro-football-reference.com/years/2019/passing_advanced.htm"
player_std_stats_url <- "https://www.pro-football-reference.com/years/2019/passing.htm"

adv_stats_html <- read_html(player_adv_stats_url)
std_stats_html <- read_html(player_std_stats_url)
```
Scrape and prepare standard passing data
```{r scrape_ay}
std_pass_df <- std_stats_html %>%
  html_node("#div_passing") %>%
  html_node("table") %>%
  html_table()

names(std_pass_df)[26] <- "Yds_sacks"

# Keep only the top 29 QBs
std_pass_df$Rk <- as.numeric(as.character(std_pass_df$Rk))
std_pass_df <- std_pass_df %>%
  filter(Rk <= 29)

# Convert all columns to numeric
std_pass_df$Age <- as.numeric(as.character(std_pass_df$Age))
std_pass_df$G <- as.numeric(as.character(std_pass_df$G))
std_pass_df$GS <- as.numeric(as.character(std_pass_df$GS))
std_pass_df$Cmp <- as.numeric(as.character(std_pass_df$Cmp))
std_pass_df$Att <- as.numeric(as.character(std_pass_df$Att))
std_pass_df$'Cmp%' <- as.numeric(sub('%', '', as.character(std_pass_df$'Cmp%')))/100
std_pass_df$Yds <- as.numeric(as.character(std_pass_df$Yds))
std_pass_df$TD <- as.numeric(as.character(std_pass_df$TD))
std_pass_df$'TD%' <- as.numeric(sub('%', '', as.character(std_pass_df$'TD%')))/100
std_pass_df$Int <- as.numeric(as.character(std_pass_df$Int))
std_pass_df$'Int%' <- as.numeric(sub('%', '', as.character(std_pass_df$'Int%')))/100
std_pass_df$'1D' <- as.numeric(as.character(std_pass_df$'1D'))
std_pass_df$Lng <- as.numeric(as.character(std_pass_df$Lng))
std_pass_df$'Y/A' <- as.numeric(as.character(std_pass_df$'Y/A'))
std_pass_df$'AY/A' <- as.numeric(as.character(std_pass_df$'AY/A'))
std_pass_df$'Y/C' <- as.numeric(as.character(std_pass_df$'Y/C'))
std_pass_df$'Y/G' <- as.numeric(as.character(std_pass_df$'Y/G'))
std_pass_df$Rate <- as.numeric(as.character(std_pass_df$Rate))
std_pass_df$QBR <- NULL
std_pass_df$Sk <- as.numeric(as.character(std_pass_df$Sk))
std_pass_df$Yds_sacks <- as.numeric(as.character(std_pass_df$Yds_sacks))
std_pass_df$'NY/A' <- as.numeric(as.character(std_pass_df$'NY/A'))
std_pass_df$'ANY/A' <- as.numeric(as.character(std_pass_df$'ANY/A'))
std_pass_df$'Sk%' <- as.numeric(sub('%', '', as.character(std_pass_df$'Sk%')))/100
std_pass_df$'4QC' <- as.numeric(as.character(std_pass_df$'4QC'))
std_pass_df$GWD <- as.numeric(as.character(std_pass_df$GWD))

std_pass_df[is.na(std_pass_df)] <- 0
```

Make all variables standardized per game played and remove the old non-standardized variables
```{r standardize}
std_pass_df <- std_pass_df %>%
  mutate('Cmp/G' = Cmp/G) %>%
  mutate('Att/G' = Att/G) %>%
  mutate('Yds/G' = Yds/G) %>%
  mutate('TD/G' = TD/G) %>%
  mutate('Int/G' = Int/G) %>%
  mutate('1D/G' = std_pass_df$'1D'/G) %>%
  mutate('Sk/G' = Sk/G) %>%
  mutate('Yds_sacks/G' = Yds_sacks/G) %>%
  mutate('4QC/G' = std_pass_df$'4QC'/G) %>%
  mutate('GWD/G' = GWD/G)

std_pass_df$Cmp <- NULL
std_pass_df$Att <- NULL
std_pass_df$Yds <- NULL
std_pass_df$TD <- NULL
std_pass_df$Int <- NULL
std_pass_df$'1D' <- NULL
std_pass_df$Sk <- NULL
std_pass_df$Yds_sacks <- NULL
std_pass_df$'4QC' <- NULL
std_pass_df$GWD <- NULL
std_pass_df$Lng <- NULL
```

Scrape the advanced passing air yards dataframe data
```{r scrape_ay}
adv_pass_ay_df <- adv_stats_html %>%
  html_node("#all_ks_passing_detailed_air_yards") %>%
  html_node("table") %>%
  html_table()

names(adv_pass_ay_df) <- as.character(unlist(adv_pass_ay_df[1,]))
adv_pass_ay_df <- adv_pass_ay_df[-1,]
```

Similarly, scrape the advanced passing accuracy, passing under pressure, and passing based on play type statistics.
```{r scrape_acc_press}
# Accuracy Data
adv_pass_acc_df <- adv_stats_html %>%
  html_node("#all_ks_passing_detailed_accuracy") %>%
  html_node("table") %>%
  html_table()

names(adv_pass_acc_df) <- as.character(unlist(adv_pass_acc_df[1,]))
adv_pass_acc_df <- adv_pass_acc_df[-1,]

# Pressure Data
adv_pass_press_df <- adv_stats_html %>%
  html_node("#all_ks_passing_detailed_pressure") %>%
  html_node("table") %>%
  html_table()

names(adv_pass_press_df) <- as.character(unlist(adv_pass_press_df[1,]))
adv_pass_press_df <- adv_pass_press_df[-1,]

# Advanced Passing Play Type
adv_pass_play_type_df <- adv_stats_html %>%
  html_node("#all_ks_passing_detailed_play_type") %>%
  html_node("table") %>%
  html_table()

names(adv_pass_play_type_df) <- as.character(unlist(adv_pass_play_type_df[1,]))
adv_pass_play_type_df <- adv_pass_play_type_df[-1,]

# Gonna need to do some renaming of the columns to distinguish between RPO and Play Action (PA) plays
names(adv_pass_play_type_df)[11] <- "RPO_Plays"
names(adv_pass_play_type_df)[12] <- "RPO_Yds"
names(adv_pass_play_type_df)[13] <- "RPO_PassAtt"
names(adv_pass_play_type_df)[14] <- "RPO_PassYds"
names(adv_pass_play_type_df)[15] <- "RPO_RushAtt"
names(adv_pass_play_type_df)[16] <- "RPO_RushYds"
names(adv_pass_play_type_df)[17] <- "PA_PassAtt"
names(adv_pass_play_type_df)[18] <- "PA_PassYds"
```

Join all of the tables together to create one conclusive advanced passing player analytics table. We'll aslo convert all of the columns that have numbers in them to numeric, and remove NA's.

```{r join_pass_stats}
adv_passing_df <- adv_pass_acc_df %>%
  inner_join(adv_pass_ay_df, by=c("Rk","Player","Tm","Age","Pos", "G", "GS", "Cmp", "Att", "Yds")) %>%
  inner_join(adv_pass_play_type_df, by=c("Rk","Player","Tm","Age","Pos", "G", "GS", "Cmp", "Att", "Yds")) %>%
  inner_join(adv_pass_press_df, by=c("Rk","Player","Tm","Age","Pos", "G", "GS", "Cmp", "Att", "Yds"))

adv_passing_df$GS <- as.numeric(as.character(adv_passing_df$GS))
adv_passing_df$Age <- as.numeric(as.character(adv_passing_df$Age))
adv_passing_df$G <- as.numeric(as.character(adv_passing_df$G))
adv_passing_df$Cmp <- as.numeric(as.character(adv_passing_df$Cmp))
adv_passing_df$Att <- as.numeric(as.character(adv_passing_df$Att))
adv_passing_df$Yds <- as.numeric(as.character(adv_passing_df$Yds))
adv_passing_df$Bats <- as.numeric(as.character(adv_passing_df$Bats))
adv_passing_df$ThAwy <- as.numeric(as.character(adv_passing_df$ThAwy))
adv_passing_df$Spikes <- as.numeric(as.character(adv_passing_df$Spikes))
adv_passing_df$Drops <- as.numeric(as.character(adv_passing_df$Drops))
adv_passing_df$'Drop%' <- as.numeric(sub('%', '', as.character(adv_passing_df$'Drop%')))/100
adv_passing_df$BadTh <- as.numeric(as.character(adv_passing_df$BadTh))
adv_passing_df$'Bad%' <- as.numeric(sub('%', '', as.character(adv_passing_df$'Bad%')))/100
adv_passing_df$OnTgt <- as.numeric(as.character(adv_passing_df$OnTgt))
adv_passing_df$'OnTgt%' <- as.numeric(sub('%', '', as.character(adv_passing_df$'OnTgt%')))/100
adv_passing_df$IAY <- as.numeric(as.character(adv_passing_df$IAY))
adv_passing_df$'IAY/PA' <- as.numeric(as.character(adv_passing_df$'IAY/PA'))
adv_passing_df$CAY <- as.numeric(as.character(adv_passing_df$CAY))
adv_passing_df$'CAY/Cmp' <- as.numeric(as.character(adv_passing_df$'CAY/Cmp'))
adv_passing_df$'CAY/PA' <- as.numeric(as.character(adv_passing_df$'CAY/PA'))
adv_passing_df$YAC <- as.numeric(as.character(adv_passing_df$YAC))
adv_passing_df$'YAC/Cmp' <- as.numeric(as.character(adv_passing_df$'YAC/Cmp'))
adv_passing_df$RPO_Plays <- as.numeric(as.character(adv_passing_df$RPO_Plays))
adv_passing_df$RPO_Yds <- as.numeric(as.character(adv_passing_df$RPO_Yds))
adv_passing_df$RPO_PassAtt <- as.numeric(as.character(adv_passing_df$RPO_PassAtt))
adv_passing_df$RPO_PassYds <- as.numeric(as.character(adv_passing_df$RPO_PassYds))
adv_passing_df$RPO_RushAtt <- as.numeric(as.character(adv_passing_df$RPO_RushAtt))
adv_passing_df$RPO_RushYds <- as.numeric(as.character(adv_passing_df$RPO_RushYds))
adv_passing_df$PA_PassAtt <- as.numeric(as.character(adv_passing_df$PA_PassAtt))
adv_passing_df$PA_PassYds <- as.numeric(as.character(adv_passing_df$PA_PassYds))
adv_passing_df$Sk <- as.numeric(as.character(adv_passing_df$Sk))
adv_passing_df$PktTime <- as.numeric(as.character(adv_passing_df$PktTime))
adv_passing_df$Bltz <- as.numeric(as.character(adv_passing_df$Bltz))
adv_passing_df$Hrry <- as.numeric(as.character(adv_passing_df$Hrry))
adv_passing_df$Hits <- as.numeric(as.character(adv_passing_df$Hits))
adv_passing_df$Scrm <- as.numeric(as.character(adv_passing_df$Scrm))
adv_passing_df$'Yds/Scr' <- as.numeric(as.character(adv_passing_df$'Yds/Scr'))

adv_passing_df[is.na(adv_passing_df)] <- 0

adv_passing_df$Rk <- as.numeric(as.character(adv_passing_df$Rk))
adv_passing_df <- adv_passing_df %>%
  filter(Rk <= 29)

head(adv_passing_df)
```

Standardize the advanced passing stats like we did before with the standard passing data.
```{r standardize_adv}

rm(adv_pass_acc_df, adv_pass_ay_df, adv_pass_play_type_df, adv_pass_press_df)

adv_passing_df <- adv_passing_df %>%
  mutate('Bats/G' = Bats/G) %>%
  mutate('ThAwy/G' = ThAwy/G) %>%
  mutate('RPO_Plays/G' = RPO_Plays/G) %>%
  mutate('RPO_Yds/G' = RPO_Yds/G) %>%
  mutate('RPO_PassAtt/G' = RPO_PassAtt/G) %>%
  mutate('RPO_PassYds/G' = RPO_PassYds/G) %>%
  mutate('RPO_RushAtt/G' = RPO_RushAtt/G) %>%
  mutate('RPO_RushYds/G' = RPO_RushYds/G) %>%
  mutate('PA_PassAtt/G' = PA_PassAtt/G) %>%
  mutate('PA_PassYds/G' = PA_PassYds/G) %>%
  mutate('Bltz/G' = Bltz/G) %>%
  mutate('Hrry/G' = Hrry/G) %>%
  mutate('Hits/G' = Hits/G) %>%
  mutate('Scrm/G' = Scrm/G)

adv_passing_df$Cmp <- NULL
adv_passing_df$Att <- NULL
adv_passing_df$Yds <- NULL
adv_passing_df$Spikes <- NULL
adv_passing_df$IAY <- NULL
adv_passing_df$CAY <- NULL
adv_passing_df$YAC <- NULL
adv_passing_df$Sk <- NULL
adv_passing_df$Bats <- NULL
adv_passing_df$ThAwy <- NULL
adv_passing_df$Drops <- NULL
adv_passing_df$BadTh <- NULL
adv_passing_df$OnTgt <- NULL
adv_passing_df$IAY <- NULL
adv_passing_df$CAY <- NULL
adv_passing_df$YAC <- NULL
adv_passing_df$Drops <- NULL
adv_passing_df$RPO_Plays <- NULL
adv_passing_df$RPO_Yds <- NULL
adv_passing_df$RPO_PassAtt <- NULL
adv_passing_df$RPO_PassYds <- NULL
adv_passing_df$RPO_RushAtt <- NULL
adv_passing_df$RPO_RushYds <- NULL
adv_passing_df$PA_PassAtt <- NULL
adv_passing_df$PA_PassYds <- NULL
adv_passing_df$Bltz <- NULL
adv_passing_df$Hrry <- NULL
adv_passing_df$Hits <- NULL
adv_passing_df$Scrm <- NULL
```

Now we'll join the standardized datasets (of standard and advanced passing data) to get one super passing dataset.
```{r join_std_adv}
pass_df <- std_pass_df %>%
  inner_join(adv_passing_df, by=c("Rk","Player","Tm","Age","Pos", "G", "GS"))

```

Now lets convert the QB records into a useful statstic... win percentage.
```{r win_%}

pass_df$QBrec <- as.numeric(gsub( "-.*$", "", pass_df$QBrec))

pass_df <- pass_df %>%
  mutate('Win%' = QBrec/G)

pass_df$QBrec <- NULL

write.csv(pass_df, "pass_df.csv")

```
Beautiful.

### 2020-05-08 Clustering/2019_qb_rb_wr_clustering.Rmd
initial commit of files

---
title: "K-Means Clustering of 2019 NFL Quarterbacks, Running Backs and Receivers"
author: "Philip Ipe"
date: "4/3/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(tidyverse)
library(knitr)
```

## Data Preparation

We have scraped and prepared data from Pro-Football Reference of the top 29-ranked quarterbacks in the 2019 NFL season.

Insert table of all of the variables
```{r import}
pass_df <- read.csv("pass_df.csv")
pass_df$X <- NULL
```

## K-Means Clustering

Now let's do some clustering of the passing data we have collected. We will use the k-means method, which in very simple terms will cluster the data by grouping teams with similar passing statistics together with one another. But first, we need to determine exactly how many groups we want to cluster them into---or k. We will use the elbow method to determine our k value, for more info on the elbow method visit the interwebs.

Determining K using Elbow Method
```{r elbow}
set.seed(20)
rng<-2:18
tries <-100 #Run the K Means algorithm 100 times
avg.totw.ss <-integer(length(rng)) #Set up an empty vector to hold all of points
for(v in rng){ # For each value of the range variable
 v.totw.ss <-integer(tries) #Set up an empty vector to hold the 100 tries
 for(i in 1:tries){
 k.temp <-kmeans(pass_df[,8:52],centers=v) #Run kmeans
 v.totw.ss[i] <-k.temp$tot.withinss#Store the total withinss
 }
 avg.totw.ss[v-1] <-mean(v.totw.ss) #Average the 100 total withinss
}
plot(rng,avg.totw.ss,type="b", main="Total Within SS by Various K",
 ylab="Average Total Within Sum of Squares",
 xlab="Value of K")
```

We see that the point of the elbow is (kinda) at a value of 8, so let's roll with k=8 with our data. We'll attach the corresponding cluster number to each player now.
```{r k_means}
clusters <- kmeans(pass_df[,8:52], 8)
pass_df$cluster <- as.factor(clusters$cluster)
head(pass_df)
```

Use summaries of each cluster to say which each cluster is good at and what they are bad at.
```{r print_clusters}
#for (cluster_num in 1:8) {
 # adv_passing_df %>%
  #  filter(cluster ==cluster_num) %>%
   # print(Player)
#}

pass_df %>%
  group_by(cluster) %>%
  summarize_all(mean) %>%
  print()
```
From our analysis, we find that players like Jared Goff, Jameis Winston, and Lamar Jackson are in a class of their own.
Now let's plot bad throw percentage vs. on target percentage for each quarterback.

```{r plotting_bad_v_ontgt}
pass_df %>%
  group_by(cluster) %>%
  ggplot(mapping=aes(x=`Bad.`, y=`OnTgt.`)) + 
  geom_point(aes(color = cluster))
```

```{r plotting_more}
pass_df %>%
  group_by(cluster) %>%
  ggplot(mapping=aes(x=`Rate`, y=`Win.`)) + 
  geom_point(aes(color = cluster))
```

Add salaries of each player...do some clusters get paid more than others?

Do some clusters win more than others?

Add years starting to each QB, do some clusters just have less experience than others?

### 2020-05-08 Drafting/draft_analysis.Rmd
initial commit of files

---
title: "NFL Draft Analysis"
author: "Philip Ipe"
date: "4/25/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1)
library(tidyverse)
library(rvest)
setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Analytics/Drafting")
```

## Preparation

Load the NFL draft history data, containing the draft data from rounds 1-4 from the 1970-2016 NFL draft.
```{r cars}
master_df <- read.csv("rd1_4_draft.csv")
head(master_df)

master_df <- master_df[complete.cases(master_df[,12:13]),]
master_df$Rnd <- as.factor(master_df$Rnd)
```

## Trends in Average value
Has average value for players selected in different rounds gone up over time?
Have GMs gotten better at drafting in the first round, second round, etc...?

```{r avg_val_rd1}

avg_val_df <- master_df %>%
  group_by(year, Rnd) %>%
  summarise(Avg_AV = mean(CarAV))

avg_val_df %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=year, y=Avg_AV, color=Rnd)) +
    geom_point() +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for Players Drafted in Different Rounds Over Time") + 
    geom_smooth(method=lm) 
```

Seems like GMs have not gotten better at drafting in any of the first four rounds. In fact, we see a slight decline in average value for all rounds over time. We do see that GMs are selecting better players in the earlier rounds, as generally the average value for the earlier rounds is consistently higher than the lower rounds throughout the years. We see the trends spread out as time goes on as well.

## Analyzing Different Positions

```{r positions}
pos_df <- master_df %>%
  mutate(period = cut(year, breaks=4)) %>%
  group_by(period, Pos, Rnd) %>%
  summarise(Avg_AV = mean(CarAV))

master_df2 <- master_df

master_df2$Rnd <- as.numeric(master_df2$Rnd)
```

Lets analyze the average values of different positions grouped by round and time period in which they were drafted. Let's start with the most important position in football.

```{r qb}
pos_df %>%
  filter(Pos == "QB") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for QBs Drafted in Different Rounds Over Different Time Periods")
```
```{r qb2}

master_df2 %>%
  filter(year >= 2000) %>%
  filter(Pos == "QB") %>%
  ggplot(mapping=aes(x=Pick, y=CarAV)) +
    geom_point()  +
    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for QBs Drafted in the 2000s") + 
    geom_smooth(method='loess')
```

```{r qb3}

master_df2 %>%
  filter(year >= 2000) %>%
  filter(Pos == "QB") %>%
  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
    geom_violin()  +
    labs(x = "Rnd", y = "Career Value", title = "Average Value over picks for QBs Drafted in the 2000s") + 
    geom_smooth(method=lm)
```

In the past 20 years (2000-2020), it is much more likely to hit on a QB in the first round than in any of the other rounds. We see a huge drop in average value after the first round in the past 20 years. Average value dips in half after that, and continues to decrease until the 7th round. It is interesting to note that the 7th round averge value shoots up again to about the same average values of the 3rd and 4th round in the 2000s.

Before then, we see that the 6th round was a  value pick for QB, although Tom Brady might be the outlier that is skewing those numbers.

Now let's analyze the second most important position...
```{r de}
pos_df %>%
  filter(Pos == "DE") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for DEs Drafted in Different Rounds Over Different Time Periods")
```
Not as big of a dip in average value from 1-4 round as other positions.

```{r wr}
pos_df %>%
  filter(Pos == "WR") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for WRs Drafted in Different Rounds Over Different Time Periods")
```
Rounds 4-6 have around the same value, suggests 5th and 6th rounds are value picks for WR. We also don't see as huge of a drop off in value from 1st to 2nd round here (and even from 2nd to 3rd) as we did when we looked at the quarterback position.

```{r rb}
pos_df %>%
  filter(Pos == "RB") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for RBs Drafted in Different Rounds Over Different Time Periods")
```
Big drop in average value after first round, almost half (not as much as WR). Pretty surprising given that teams are looking to draft RBs in later rounds nowadays.
```{r ot}
pos_df %>%
  filter(Pos == "T") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for Ts Drafted in Different Rounds Over Different Time Periods")
```
We don't see as big a drop in value from first-fourth rounds as wel do in other positions in the 2000s. Suggests there is typically more value in second round-fourth round T picks than other positions.

```{r g}
pos_df %>%
  filter(Pos == "G") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for Gs Drafted in Different Rounds Over Different Time Periods")
```

```{r DB}
pos_df %>%
  filter(Pos == "DB") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for DBs Drafted in Different Rounds Over Different Time Periods")
```
In the 2000s, value for DBs picked in rounds 3-7 have about the same value.

```{r TE}
pos_df %>%
  filter(Pos == "TE") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for TEs Drafted in Different Rounds Over Different Time Periods")
```
Are there any positions that particularly suck when drafted after a certain round? 
What position show the least and most correlation between the pick number and average value brought?
Which positions have been the best value picks?

### 2020-05-08 Drafting/qb_draft_analysis.Rmd
initial commit of files

---
title: 'NFL Draft Analysis: The Quarterback'
author: "Philip Ipe"
date: "5/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Analytics/Drafting")
```

# Part 1: Data Collection

We will scrape historical draft data from Pro-Football Reference from years 1970 till 2016. We will filter out any selections after round 7, as since 1994, the NFL draft has only consisted of 7 rounds.

```{r scraping}

master_df <- data.frame()
for (year in 1970:2016){

  draft_url <- paste("https://www.pro-football-reference.com/years/", year, "/draft.htm", sep="")
  draft_html <- read_html(draft_url)

  df <- draft_html %>%
    html_node("table") %>%
    html_table()

  names(df) <- as.character(unlist(df[1,]))
  df <- df[-1,]

  # get rid of the columns we dont care about
  df <- df[,1:13]

  df$Rnd <- as.integer(df$Rnd)
  df$Age <- as.integer(df$Age)
  df$CarAV <- as.integer(df$CarAV)
  df$DrAV <- as.integer(df$DrAV)
  df$G <- as.integer(df$G)

  # filter out certain rounds
  df <- df %>%
    filter(Rnd < 8)

  # get rid of blank rows
  df <-df[!(df$Player=="Player"),]

  # add year drafted to each observation
  df <- df %>%
    mutate(year = year)

  master_df <- rbind(master_df, df)
}
```
# Part 2: Data Processing and Preparation

Load the NFL draft history data, containing the draft data from rounds 1-4 from the 1970-2016 NFL draft.
```{r cars}
head(master_df)

master_df <- master_df[complete.cases(master_df[,12:13]),]
master_df$Rnd <- as.factor(master_df$Rnd)

master_df$Age <- as.integer(master_df$Age)
master_df$CarAV <- as.integer(master_df$CarAV)
master_df$DrAV <- as.integer(master_df$DrAV)
master_df$G <- as.integer(master_df$G)
master_df$Pick <- as.integer(master_df$Pick)
```

First, we want to break up our data into 4 periods. Then for each period, we will take average career value of each position grouped by each round that the player was selected. We will make a new pos_df dataframe with these groupings.

```{r positions}
pos_df <- master_df %>%
  mutate(period = cut(year, breaks=4)) %>%
  group_by(period, Pos, Rnd) %>%
  summarise(Avg_AV = mean(CarAV))

head(pos_df)
```

# Part 3: Exploratory Data Analysis

## Trends in Average value
Has average value for players selected in different rounds gone up over time?
Have GMs gotten better at drafting in the first round, second round, etc...?

```{r avg_val_rd1}

avg_val_df <- master_df %>%
  group_by(year, Rnd) %>%
  summarise(Avg_AV = mean(CarAV))

avg_val_df %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=year, y=Avg_AV, color=Rnd)) +
    geom_point() +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for Players Drafted in Different Rounds Over Time") + 
    geom_smooth(method=lm) 
```

Seems like GMs have not gotten better at drafting in any of the first four rounds. In fact, we see a slight decline in average value for all rounds over time. We do see that GMs are selecting better players in the earlier rounds, as generally the average value for the earlier rounds is consistently higher than the lower rounds throughout the years. We see the trends spread out as time goes on as well.

## Analyzing the Quarterback position

Lets analyze the average values of different positions grouped by round and time period in which they were drafted. Let's start with the most important position in football.

```{r qb}
pos_df %>%
  filter(Pos == "QB") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for QBs Drafted in Different Rounds Over Different Time Periods")
```
In the past 20 years (2000-2020), it is much more likely to hit on a QB in the first round than in any of the other rounds. We see a huge drop in average value after the first round in the past 20 years. Average value dips in half after that, and continues to decrease until the 7th round. It is interesting to note that the 7th round averge value shoots up again to about the same average values of the 3rd and 4th round in the 2000s.

Before then, from 1990-2000, we see that the 6th round was a  value pick for QB, although Tom Brady might be the outlier that is skewing those numbers. But we also see that there was far more value in the second round QB selection from 1980-2000 than in the past 20 years.

```{r qb2}

master_df %>%
  filter(year >= 2000) %>%
  filter(Pos == "QB") %>%
  ggplot(mapping=aes(x=Pick, y=CarAV)) +
    geom_point()  +
    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for QBs Drafted in the 2000s") + 
    geom_smooth(method='loess')
```

As we would expect the trend is that there is a decline in value until about pick 100, where the career value seems to be about the same from about pick 100-250. This would suggest that seelcting a QB at picks 100-150 would, statistically speaking, provide less bang for your buck, as you could likely get a player of the same quality with some of your later draft capital.

```{r qb3}

master_df %>%
  filter(year >= 2000) %>%
  filter(Pos == "QB") %>%
  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
    geom_violin()  +
    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for QBs Drafted in the 2000s") + 
    geom_smooth(method=lm)
```

As we would expect, the violin plots show us that most of the QBs selected in rounds 5-7 turn out to be busts, as most of the career values end up towards the bottom of the violin.

However, the distributions reaffirm what we have seen before in our Round 1 QBs. The spread of the data is far greater in round 1, and decreases heavily in round 2, and from there the decreases in spread are less drastic, suggesting that you have more of a shot at getting a good player in round 1 than in any other round.

## Analyzing the WR and RB positions

Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft.

# Part 4: Machine Learning

## Predicing the Rookie Offensive MVP

### 2020-05-08 Drafting/scrape_draft_history.R

@@ -1,0 +1,46 @@
+library(tidyverse)
+library(rvest)
+
+setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Analytics/Drafting")
+master_df <- data.frame()
+for (year in 1970:2016){
+  
+  draft_url <- paste("https://www.pro-football-reference.com/years/", year, "/draft.htm", sep="")
+  draft_html <- read_html(draft_url)
+
+  df <- draft_html %>%
+    html_node("table") %>%
+    html_table()
+
+  names(df) <- as.character(unlist(df[1,]))
+  df <- df[-1,]
+  
+  # get rid of the columns we dont care about
+  df <- df[,1:13]
+  
+  df$Rnd <- as.integer(df$Rnd)
+  df$Age <- as.integer(df$Age)
+  df$CarAV <- as.integer(df$CarAV)
+  df$DrAV <- as.integer(df$DrAV)
+  df$G <- as.integer(df$G)
+  
+  # filter out certain rounds
+  df <- df %>%
+    filter(Rnd < 8)
+  
+  # get rid of blank rows
+  df <-df[!(df$Player=="Player"),]
+  
+  # add year drafted to each observation
+  df <- df %>%
+    mutate(year = year)
+
+  master_df <- rbind(master_df, df)
+}
+
+master_df$Rnd <- as.integer(master_df$Rnd)
+master_df$Age <- as.integer(master_df$Age)
+master_df$CarAV <- as.integer(master_df$CarAV)
+master_df$DrAV <- as.integer(master_df$DrAV)
+master_df$G <- as.integer(master_df$G)
+write.csv(master_df, "rd1_4_draft.csv")

### 2020-05-08 Drafting/qb_draft_analysis.Rmd
Scrape Offensive ROTY data and disable warnings on knit

@@
 knitr::opts_chunk$set(echo = TRUE)
+options(warn = -1)
 library(tidyverse)
@@
 }
+```
+
+Next we will scrape the AP Offensive Rookie of the Year Winners from the same website.
+
+```{r scrape_ROTY}
+
+roty_url <- "https://www.pro-football-reference.com/awards/ap-offensive-rookie-of-the-year-award.htm"
+draft_html <- read_html(roty_url)
+
+roty_df <- draft_html %>%
+  html_node("table") %>%
+  html_table()
 ```

### 2020-05-08 Drafting/qb_draft_analysis.Rmd
Add WR and RB plots and analysis

@@
 ---
-title: 'NFL Draft Analysis: The Quarterback'
+title: 'NFL Draft Analysis'
 author: "Philip Ipe"
@@
 library(rvest)
+library(gridExtra)
 setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Analytics/Drafting")
@@
 ```{r qb}
-pos_df %>%
+qb_bar <- pos_df %>%
   filter(Pos == "QB") %>%
@@
     labs(x = "Year", y = "Average Career Value", title = "Average Value for QBs Drafted in Different Rounds Over Different Time Periods")
+
+qb_bar
 ```
+
 In the past 20 years (2000-2020), it is much more likely to hit on a QB in the first round than in any of the other rounds.
@@
 
+Now let's focus our attention on the past 20 years and graph some visualizations of that data.
 ```{r qb2}
 
-master_df %>%
+qb_av_pick <- master_df %>%
   filter(year >= 2000) %>%
@@
     geom_smooth(method='loess')
+
+qb_av_pick
 ```
@@
 
-master_df %>%
+qb_violin <- master_df %>%
   filter(year >= 2000) %>%
@@
     geom_smooth(method=lm)
+
+qb_violin
 ```
@@
 
+It is interesting to note that the general trend has been to select RBs later in the draft, as RBs have always been considered as more expendable than a QB, and there seems to be a notion that you can always get a quality back in the later rounds, and that using first round capital on a RB is a poor management decision.
+
+```{r rb_wr1}
+
+rb_bar <- pos_df %>%
+  filter(Pos == "RB") %>%
+  group_by(Rnd) %>%
+  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
+    geom_bar(stat="identity", position = "dodge") +
+    labs(x = "Year", y = "Average Career Value", title = "Average Value for RBs Drafted in Different Rounds Over Different Time Periods")
+
+wr_bar <- pos_df %>%
+  filter(Pos == "WR") %>%
+  group_by(Rnd) %>%
+  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
+    geom_bar(stat="identity", position = "dodge") +
+    labs(x = "Year", y = "Average Career Value", title = "Average Value for WRs Drafted in Different Rounds Over Different Time Periods")
+
+grid.arrange(qb_bar, rb_bar, wr_bar, ncol=3)
+```
+
+Not as drastic of a difference in rounds 1 and 2 in the WR group. The RB average value in the 2000s seems to resemble the QB position.
+
+```{r rb_wr2}
+
+rb_av_pick <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "RB") %>%
+  ggplot(mapping=aes(x=Pick, y=CarAV)) +
+    geom_point()  +
+    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for RBs") + 
+    geom_smooth(method='loess')
+
+wr_av_pick <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "WR") %>%
+  ggplot(mapping=aes(x=Pick, y=CarAV)) +
+    geom_point()  +
+    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for WRs") + 
+    geom_smooth(method='loess')
+
+grid.arrange(qb_av_pick, rb_av_pick, wr_av_pick, ncol=3)
+
+```
+
+Much less steep of a drop off between round 1 to 2 in the WR group.
+
+Again we see a trend that by pick 100 the value seems to hit a plateue.
+
+```{r rb_wr3}
+
+rb_violin <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "RB") %>%
+  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
+    geom_violin()  +
+    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for RBs") + 
+    geom_smooth(method=lm)
+
+wr_violin <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "WR") %>%
+  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
+    geom_violin()  +
+    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for WRs") + 
+    geom_smooth(method=lm)
+
+grid.arrange(qb_violin, rb_violin, wr_violin, ncol=3)
+```
+
+RBs in Round 1 always produce.
+We again see that round 2 is a value round for WRs as it appears to resemble the type of value you get out of round 1.
+
 # Part 4: Machine Learning
 
-## Predicing the Rookie Offensive MVP
+## Predicting the Rookie Offensive MVP

### 2020-05-09 Drafting/draft_analysis.Rmd
Some renaming

@@
 ---
-title: "NFL Draft Analysis"
+title: 'NFL Draft Analysis'
 author: "Philip Ipe"
-date: "4/25/2020"
-output: pdf_document
+date: "5/8/2020"
+output: html_document
+knit: (function(input_file, encoding) {
+  out_dir <- 'docs';
+  rmarkdown::render(input_file,
+ encoding=encoding,
+ output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
 ---
@@
 
-## Preparation
+# Part 1: Data Collection
+
+We will scrape historical draft data from Pro-Football Reference from years 1970 till 2016.
+We chose to stop at 2016 since rookies usually sign 4 year deals, so the four year mark is typically when each team will finish evaulating each player they drafted, and make a decision on whether or not to resign the player.
+So, the jury is still out on players drafted after 2016.
+We will filter out any selections after round 7, as since 1994, the NFL draft has only consisted of 7 rounds.
+
+```{r scraping}
+
+master_df <- data.frame()
+for (year in 1970:2016){
+
+  draft_url <- paste("https://www.pro-football-reference.com/years/", year, "/draft.htm", sep="")
+  draft_html <- read_html(draft_url)
+
+  df <- draft_html %>%
+    html_node("table") %>%
+    html_table()
+
+  names(df) <- as.character(unlist(df[1,]))
+  df <- df[-1,]
+
+  # get rid of the columns we dont care about
+  df <- df[,1:13]
+
+  df$Rnd <- as.integer(df$Rnd)
+  df$Age <- as.integer(df$Age)
+  df$CarAV <- as.integer(df$CarAV)
+  df$DrAV <- as.integer(df$DrAV)
+  df$G <- as.integer(df$G)
+
+  # filter out certain rounds
+  df <- df %>%
+    filter(Rnd < 8)
+
+  # get rid of blank rows
+  df <-df[!(df$Player=="Player"),]
+
+  # add year drafted to each observation
+  df <- df %>%
+    mutate(year = year)
+
+  master_df <- rbind(master_df, df)
+}
+```
+
+Next we will scrape the AP Offensive Rookie of the Year Winners from the same website.
+
+```{r scrape_ROTY}
+
+roty_url <- "https://www.pro-football-reference.com/awards/ap-offensive-rookie-of-the-year-award.htm"
+draft_html <- read_html(roty_url)
+
+roty_df <- draft_html %>%
+  html_node("table") %>%
+  html_table()
+```
+# Part 2: Data Processing and Preparation
 
@@
 ```{r cars}
-master_df <- read.csv("rd1_4_draft.csv")
 head(master_df)
@@
 master_df$Rnd <- as.factor(master_df$Rnd)
+
+master_df$Age <- as.integer(master_df$Age)
+master_df$CarAV <- as.integer(master_df$CarAV)
+master_df$DrAV <- as.integer(master_df$DrAV)
+master_df$G <- as.integer(master_df$G)
+master_df$Pick <- as.integer(master_df$Pick)
 ```
+
+First, we want to break up our data into 4 periods.
+Then for each period, we will take average career value of each position grouped by each round that the player was selected.
+We will make a new pos_df dataframe with these groupings.
+
+```{r positions}
+pos_df <- master_df %>%
+  mutate(period = cut(year, breaks=4)) %>%
+  group_by(period, Pos, Rnd) %>%
+  summarise(Avg_AV = mean(CarAV))
+
+head(pos_df)
+```
+
+# Part 3: Exploratory Data Analysis
 
@@
 
-## Analyzing Different Positions
-
-```{r positions}
-pos_df <- master_df %>%
-  mutate(period = cut(year, breaks=4)) %>%
-  group_by(period, Pos, Rnd) %>%
-  summarise(Avg_AV = mean(CarAV))
-
-master_df2 <- master_df
-
-master_df2$Rnd <- as.numeric(master_df2$Rnd)
-```
+## Analyzing the Quarterback position
 
@@
     labs(x = "Year", y = "Average Career Value", title = "Average Value for QBs Drafted in Different Rounds Over Different Time Periods")
-```
-```{r qb2}
 
-master_df2 %>%
-  filter(year >= 2000) %>%
-  filter(Pos == "QB") %>%
-  ggplot(mapping=aes(x=Pick, y=CarAV)) +
-    geom_point()  +
-    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for QBs Drafted in the 2000s") + 
-    geom_smooth(method='loess')
-```
-
-```{r qb3}
-
-master_df2 %>%
-  filter(year >= 2000) %>%
-  filter(Pos == "QB") %>%
-  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
-    geom_violin()  +
-    labs(x = "Rnd", y = "Career Value", title = "Average Value over picks for QBs Drafted in the 2000s") + 
-    geom_smooth(method=lm)
+qb_bar
 ```
@@
 
-Before then, we see that the 6th round was a  value pick for QB, although Tom Brady might be the outlier that is skewing those numbers.
+Before then, from 1990-2000, we see that the 6th round was a  value pick for QB, although Tom Brady might be the outlier that is skewing those numbers.
+But we also see that there was far more value in the second round QB selection from 1980-2000 than in the past 20 years.
 
-Now let's analyze the second most important position...
-```{r de}
-pos_df %>%
-  filter(Pos == "DE") %>%
+Now let's focus our attention on the past 20 years and graph some visualizations of that data.
+```{r qb2}
+
+qb_av_pick <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "QB") %>%
+  ggplot(mapping=aes(x=Pick, y=CarAV)) +
+    geom_point()  +
+    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for QBs Drafted in the 2000s") + 
+    geom_smooth(method='loess')
+
+qb_av_pick
+```
+
+As we would expect the trend is that there is a decline in value until about pick 100, where the career value seems to be about the same from about pick 100-250.
+This would suggest that seelcting a QB at picks 100-150 would, statistically speaking, provide less bang for your buck, as you could likely get a player of the same quality with some of your later draft capital.
+
+```{r qb3}
+
+qb_violin <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "QB") %>%
+  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
+    geom_violin()  +
+    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for QBs Drafted in the 2000s") + 
+    geom_smooth(method=lm)
+
+qb_violin
+```
+
+As we would expect, the violin plots show us that most of the QBs selected in rounds 5-7 turn out to be busts, as most of the career values end up towards the bottom of the violin.
+
+However, the distributions reaffirm what we have seen before in our Round 1 QBs.
+The spread of the data is far greater in round 1, and decreases heavily in round 2, and from there the decreases in spread are less drastic, suggesting that you have more of a shot at getting a good player in round 1 than in any other round.
+
+## Analyzing the WR and RB positions
+
+Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft.
+
+It is interesting to note that the general trend has been to select RBs later in the draft, as RBs have always been considered as more expendable than a QB, and there seems to be a notion that you can always get a quality back in the later rounds, and that using first round capital on a RB is a poor management decision.
+
+```{r rb_wr1}
+
+rb_bar <- pos_df %>%
+  filter(Pos == "RB") %>%
   group_by(Rnd) %>%
@@
     geom_bar(stat="identity", position = "dodge") +
-    labs(x = "Year", y = "Average Career Value", title = "Average Value for DEs Drafted in Different Rounds Over Different Time Periods")
-```
-Not as big of a dip in average value from 1-4 round as other positions.
+    labs(x = "Year", y = "Average Career Value", title = "Average Value for RBs Drafted in Different Rounds Over Different Time Periods")
 
-```{r wr}
-pos_df %>%
+wr_bar <- pos_df %>%
   filter(Pos == "WR") %>%
@@
     labs(x = "Year", y = "Average Career Value", title = "Average Value for WRs Drafted in Different Rounds Over Different Time Periods")
-```
-Rounds 4-6 have around the same value, suggests 5th and 6th rounds are value picks for WR.
-We also don't see as huge of a drop off in value from 1st to 2nd round here (and even from 2nd to 3rd) as we did when we looked at the quarterback position.
 
-```{r rb}
-pos_df %>%
-  filter(Pos == "RB") %>%
-  group_by(Rnd) %>%
-  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
-    geom_bar(stat="identity", position = "dodge") +
-    labs(x = "Year", y = "Average Career Value", title = "Average Value for RBs Drafted in Different Rounds Over Different Time Periods")
-```
-Big drop in average value after first round, almost half (not as much as WR). Pretty surprising given that teams are looking to draft RBs in later rounds nowadays.
-```{r ot}
-pos_df %>%
-  filter(Pos == "T") %>%
-  group_by(Rnd) %>%
-  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
-    geom_bar(stat="identity", position = "dodge") +
-    labs(x = "Year", y = "Average Career Value", title = "Average Value for Ts Drafted in Different Rounds Over Different Time Periods")
-```
-We don't see as big a drop in value from first-fourth rounds as wel do in other positions in the 2000s.
-Suggests there is typically more value in second round-fourth round T picks than other positions.
-
-```{r g}
-pos_df %>%
-  filter(Pos == "G") %>%
-  group_by(Rnd) %>%
-  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
-    geom_bar(stat="identity", position = "dodge") +
-    labs(x = "Year", y = "Average Career Value", title = "Average Value for Gs Drafted in Different Rounds Over Different Time Periods")
+grid.arrange(qb_bar, rb_bar, wr_bar, ncol=3)
 ```
 
-```{r DB}
-pos_df %>%
-  filter(Pos == "DB") %>%
-  group_by(Rnd) %>%
-  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
-    geom_bar(stat="identity", position = "dodge") +
-    labs(x = "Year", y = "Average Career Value", title = "Average Value for DBs Drafted in Different Rounds Over Different Time Periods")
+Not as drastic of a difference in rounds 1 and 2 in the WR group. The RB average value in the 2000s seems to resemble the QB position.
+
+```{r rb_wr2}
+
+rb_av_pick <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "RB") %>%
+  ggplot(mapping=aes(x=Pick, y=CarAV)) +
+    geom_point()  +
+    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for RBs") + 
+    geom_smooth(method='loess')
+
+wr_av_pick <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "WR") %>%
+  ggplot(mapping=aes(x=Pick, y=CarAV)) +
+    geom_point()  +
+    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for WRs") + 
+    geom_smooth(method='loess')
+
+grid.arrange(qb_av_pick, rb_av_pick, wr_av_pick, ncol=3)
+
 ```
-In the 2000s, value for DBs picked in rounds 3-7 have about the same value.
 
-```{r TE}
-pos_df %>%
-  filter(Pos == "TE") %>%
-  group_by(Rnd) %>%
-  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
-    geom_bar(stat="identity", position = "dodge") +
-    labs(x = "Year", y = "Average Career Value", title = "Average Value for TEs Drafted in Different Rounds Over Different Time Periods")
+Much less steep of a drop off between round 1 to 2 in the WR group.
+
+Again we see a trend that by pick 100 the value seems to hit a plateue.
+
+```{r rb_wr3}
+
+rb_violin <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "RB") %>%
+  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
+    geom_violin()  +
+    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for RBs") + 
+    geom_smooth(method=lm)
+
+wr_violin <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "WR") %>%
+  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
+    geom_violin()  +
+    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for WRs") + 
+    geom_smooth(method=lm)
+
+grid.arrange(qb_violin, rb_violin, wr_violin, ncol=3)
 ```
-Are there any positions that particularly suck when drafted after a certain round? 
-What position show the least and most correlation between the pick number and average value brought?
-Which positions have been the best value picks?
+
+RBs in Round 1 always produce.
+We again see that round 2 is a value round for WRs as it appears to resemble the type of value you get out of round 1.
+
+# Part 4: Machine Learning
+
+Let's add a variable that is TRUE when the player that was drafted was Offensive Rookie of the Year, and FALSE otherwise.
+```{r add_roty}
+master_df <- master_df %>%
+  mutate(roty = Player %in% roty_df$Player)
+```
+
+## Predicting the Rookie Offensive MVP

### 2020-05-10 draft_analysis.Rmd
Make this repository work with Github Pages - remove all of the qb clustering project work - collapse all of the code into the main directory

--- 2020-05-09 Drafting/draft_analysis.Rmd
@@
 library(gridExtra)
-setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Analytics/Drafting")
+setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Draft Analytics/")
 ```

### 2020-05-10 draft_analysis.Rmd
Making Github Pages happy - moving index.html file into parent dir

@@
 knit: (function(input_file, encoding) {
-  out_dir <- 'docs';
   rmarkdown::render(input_file,
  encoding=encoding,
- output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
+ output_file=file.path(dirname(input_file), 'index.html'))})
 ---

### 2020-05-10 draft_analysis.Rmd
Add code to: - Scrape 2020 draft data - Train ML model and predict 2020 rookie offensive MVP

@@
 
+We will also scrape the data from the 2020 draft for use in our machine learning problem and save it to our df variable.
+
+```{r scrape2020}
+
+draft_url <- "https://www.pro-football-reference.com/years/2020/draft.htm"
+draft_html <- read_html(draft_url)
+
+df <- draft_html %>%
+  html_node("table") %>%
+  html_table()
+
+names(df) <- as.character(unlist(df[1,]))
+df <- df[-1,]
+
+# get rid of the columns we dont care about
+df <- df[,1:6]
+
+# filter out certain rounds
+df <- df %>%
+  filter(Rnd < 8)
+
+# get rid of blank rows
+df <-df[!(df$Player=="Player"),]
+
+# add year drafted to each observation
+df <- df %>%
+  mutate(year = year)
+
+df[is.na(df)] <- 22
+
+df$Pick <- as.integer(df$Pick)
+df$Age <- as.integer(df$Age)
+
+df$Tm <- as.factor(df$Tm)
+df$Pos <- as.factor(df$Pos)
+
+```
+
 Next we will scrape the AP Offensive Rookie of the Year Winners from the same website.
@@
 
-Load the NFL draft history data, containing the draft data from rounds 1-4 from the 1970-2016 NFL draft.
+Load the NFL draft history data, containing the draft data from rounds 1-4 from the 1970-2016 NFL draft. Remove all of the missing data entries, and convert the necessary columns to integers.
 ```{r cars}
@@
 master_df$Rnd <- as.factor(master_df$Rnd)
+master_df$Tm <- as.factor(master_df$Tm)
+master_df$Pos <- as.factor(master_df$Pos)
 
@@
 master_df <- master_df %>%
-  mutate(roty = Player %in% roty_df$Player)
+  mutate(roty = ifelse(Player %in% roty_df$Player, "Yes", "No"))
+
+master_df$roty <- as.factor(master_df$roty)
 ```
@@
 ## Predicting the Rookie Offensive MVP
+
+First, lets control the resampling of our data.
+In this case, we’re going to cross-validate the data 3 times, therefore training it 3 times on different portions of the data before settling on the best tuning parameters (for gbm it is trees, shrinkage, and interaction depth).
+
+```{r ml_control}
+
+library(caret)
+
+outcome_name <- 'roty'
+predictor_names <- c('Tm', 'Pos', 'Age', 'Pick')
+
+objControl <- trainControl(method='cv', number=3, returnResamp='none', summaryFunction = twoClassSummary, classProbs = TRUE)
+
+```
+
+Now to actually train the model...
+
+```{r train_model}
+
+model <- train(master_df[,predictor_names], master_df[,outcome_name], 
+               method = "gbm",
+               trControl = objControl,
+               metric = "ROC",
+               preProc = c("center", "scale"))
+```
+
+```{r sum_model}
+summary(model)
+```
+
+The summaries tell us how much influence each variable had on the model.
+It was interesting to see the little influence age had in predicting the model.
+It was also interesting that position had a bigger influence than the team and age.
+
+Now let's use the model to predict the probabilities that each player from the 2020 draft will become Offensive Rookie of the Year. 
+```{r eval}
+
+predictions <- predict(object=model, df[,predictor_names], type='prob')
+head(predictions)
+```

### 2020-05-11 draft_analysis.Rmd
Merge predictions with the 2020 draft data Get the top 5 predictions Some reformatting

@@
 
-df <- draft_html %>%
+df_2020 <- draft_html %>%
   html_node("table") %>%
@@
 
-names(df) <- as.character(unlist(df[1,]))
-df <- df[-1,]
-
-# get rid of the columns we dont care about
-df <- df[,1:6]
-
-# filter out certain rounds
-df <- df %>%
-  filter(Rnd < 8)
-
-# get rid of blank rows
-df <-df[!(df$Player=="Player"),]
-
-# add year drafted to each observation
-df <- df %>%
-  mutate(year = year)
-
-df[is.na(df)] <- 22
-
-df$Pick <- as.integer(df$Pick)
-df$Age <- as.integer(df$Age)
-
-df$Tm <- as.factor(df$Tm)
-df$Pos <- as.factor(df$Pos)
-
+names(df_2020) <- as.character(unlist(df_2020[1,]))
+df_2020 <- df_2020[-1,]
 ```
@@
 ```
+
 # Part 2: Data Processing and Preparation
 
-Load the NFL draft history data, containing the draft data from rounds 1-4 from the 1970-2016 NFL draft. Remove all of the missing data entries, and convert the necessary columns to integers.
+Load the NFL draft history data, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft. Remove all of the missing data entries, and convert the necessary columns to integers. 
 ```{r cars}
@@
 
+Now let's process our 2020 Draft data.
+First, we will get rid of columns we don't care about, then remove blank rows, add year drafted to each observation, and convert the necessary columns to integers, all like we did for the historical draft dataset.
+We will also filter the data to only include quarterbacks, wide receivers, and running backs.
+
+```{r process2020}
+# get rid of the columns we dont care about
+df_2020 <- df_2020[,1:6]
+
+# filter out certain rounds
+df_2020 <- df_2020 %>%
+  filter(Rnd < 8)
+
+# get rid of blank rows
+df_2020 <-df_2020[!(df_2020$Player=="Player"),]
+
+# add year drafted to each observation, and filter out the defensive players.
+df_2020 <- df_2020 %>%
+  mutate(year = year) %>%
+  filter(Pos == "QB" | Pos == "WR" | Pos == "RB" | Pos == "G" | Pos == "T")
+
+df_2020[is.na(df_2020)] <- 22
+
+df_2020$Pick <- as.integer(df_2020$Pick)
+df_2020$Age <- as.integer(df_2020$Age)
+df_2020$Tm <- as.factor(df_2020$Tm)
+df_2020$Pos <- as.factor(df_2020$Pos)
+```
+
+Now let's create a dataframe in which we can compare different positions that were drafted.
 First, we want to break up our data into 4 periods.
@@
 
+Let's also add a dataframe that contains a variable that is TRUE when the player that was drafted was Offensive Rookie of the Year, and FALSE otherwise.
+We will also filter out all of the defensive players in this new dataframe.
+
+```{r off_roty}
+off_df <- master_df %>%
+  mutate(roty = ifelse(Player %in% roty_df$Player, "Yes", "No")) %>%
+  filter(Pos == "QB" | Pos == "WR" | Pos == "RB" | Pos == "G" | Pos == "T")
+
+off_df$roty <- as.factor(off_df$roty)
+```
 # Part 3: Exploratory Data Analysis
@@
 
-Let's add a variable that is TRUE when the player that was drafted was Offensive Rookie of the Year, and FALSE otherwise.
-```{r add_roty}
-master_df <- master_df %>%
-  mutate(roty = ifelse(Player %in% roty_df$Player, "Yes", "No"))
+## Predicting the Rookie Offensive MVP
 
-master_df$roty <- as.factor(master_df$roty)
-```
-
-## Predicting the Rookie Offensive MVP
+Now, right off the bat, we hope that the model will predict that Joe Burrow has a higher likelihood of winning the MVP.
 
@@
 outcome_name <- 'roty'
-predictor_names <- c('Tm', 'Pos', 'Age', 'Pick')
+predictor_names <- c('Tm', 'Pos', 'Pick')
 
@@
 
+# Creating our Model
 Now to actually train the model...
@@
 
-model <- train(master_df[,predictor_names], master_df[,outcome_name], 
+set.seed(123)
+model <- train(off_df[,predictor_names], off_df[,outcome_name], 
                method = "gbm",
@@
 The summaries tell us how much influence each variable had on the model.
-It was interesting to see the little influence age had in predicting the model.
-It was also interesting that position had a bigger influence than the team and age.
+It was interesting to see that position had such little influence in predicting the model.
+What is even more interesting is that the team the player was drafted to seemed to influence the model more than the pick number itself!
+Before I had trained the model, I expected pick, position, and then team to have relative influence in that order!
+
+# Testing against 2017-2019 data
+
+# Predicting the 2020 Rookie Offensive MVP
 
 Now let's use the model to predict the probabilities that each player from the 2020 draft will become Offensive Rookie of the Year.
+This code will give us the probability that the model is sure about its choice that the observation will fall under the "No" and "Yes" category.
+
 ```{r eval}
 
-predictions <- predict(object=model, df[,predictor_names], type='prob')
+predictions <- predict(object=model, df_2020[,predictor_names], type='prob')
 head(predictions)
 ```
+
+Now let's join our prediction probabilities to the actual data, so we can look at the prediction probabilities for each player.
+
+```{r join_probabilities}
+
+df_2020 <- merge(df_2020, predictions, by = "row.names", all.x = FALSE, all.y = FALSE)
+df_2020$No.x <- NULL
+```
+
+Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year
+
+```{r top10_roty}
+
+top5_roty <- df_2020 %>%
+  top_n(5, Yes)
+
+head(top5_roty)
+```
+
+Joe Burrow is our model's top prediction to win the MVP!!!

### 2020-05-11 draft_analysis.Rmd
Only evaluate the skill positions for the ML problem

@@
   mutate(year = year) %>%
-  filter(Pos == "QB" | Pos == "WR" | Pos == "RB" | Pos == "G" | Pos == "T")
+  filter(Pos == "QB" | Pos == "WR" | Pos == "RB")
 
@@
   mutate(roty = ifelse(Player %in% roty_df$Player, "Yes", "No")) %>%
-  filter(Pos == "QB" | Pos == "WR" | Pos == "RB" | Pos == "G" | Pos == "T")
+  filter(Pos == "QB" | Pos == "WR" | Pos == "RB")
 
@@
 The summaries tell us how much influence each variable had on the model.
-It was interesting to see that position had such little influence in predicting the model.
+It was interesting to see that position had such absolutely zero influence in predicting the model.
+This tells us that quarterbacks, wide receivers, or running backs do not have a leg up on each other for the MVP award!
+
 What is even more interesting is that the team the player was drafted to seemed to influence the model more than the pick number itself!
@@
 
-Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year
+Now let's get the top 7 players that the model is most certain will be the Offensive Rookie of the Year
 
@@
 
-top5_roty <- df_2020 %>%
-  top_n(5, Yes)
+top10_roty <- df_2020 %>%
+  top_n(10, Yes)
 
-head(top5_roty)
+top10_roty
 ```
@@
 Joe Burrow is our model's top prediction to win the MVP!!!
+We also see that Cam Akers, Tua, Henry Ruggs III, Jerry Jeudy, CeDee Lamb, and Justin Jefferson also appear in the top 10.
+All of these players currently have the highest odds to win the the Offensive MVP according to OddsShark.

### 2020-05-11 draft_analysis.Rmd
Reformat the grids to add some titles Add some prose

@@
 rb_bar <- pos_df %>%
-  filter(Pos == "RB") %>%
+  filter(Pos == "RB" & period == "(2e+03,2.02e+03]") %>%
   group_by(Rnd) %>%
@@
     geom_bar(stat="identity", position = "dodge") +
-    labs(x = "Year", y = "Average Career Value", title = "Average Value for RBs Drafted in Different Rounds Over Different Time Periods")
+    labs(x = "Year", y = "Average Career Value", title = "Average Value for RBs")
 
 wr_bar <- pos_df %>%
-  filter(Pos == "WR") %>%
+  filter(Pos == "WR" & period == "(2e+03,2.02e+03]") %>%
   group_by(Rnd) %>%
@@
     geom_bar(stat="identity", position = "dodge") +
-    labs(x = "Year", y = "Average Career Value", title = "Average Value for WRs Drafted in Different Rounds Over Different Time Periods")
+    labs(x = "Year", y = "Average Career Value", title = "Average Value for WRs")
 
-grid.arrange(qb_bar, rb_bar, wr_bar, ncol=3)
+qb_bar <- pos_df %>%
+  filter(Pos == "QB" & period == "(2e+03,2.02e+03]") %>%
+  group_by(Rnd) %>%
+  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
+    geom_bar(stat="identity", position = "dodge") +
+    labs(x = "Year", y = "Average Career Value", title = "Average Value for QBs")
+
+grid.arrange(qb_bar, rb_bar, wr_bar, ncol=3, top = "Average Value in Different Rounds over the 2000s")
 ```
@@
 
-rb_av_pick <- master_df %>%
+qb_av_pick <- master_df %>%
   filter(year >= 2000) %>%
-  filter(Pos == "RB") %>%
+  filter(Pos == "QB") %>%
   ggplot(mapping=aes(x=Pick, y=CarAV)) +
     geom_point()  +
-    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for RBs") + 
+    labs(x = "Pick", y = "Career Value", title = "QB Average Value") + 
+    geom_smooth(method='loess')
+
+rb_av_pick <- master_df %>%
+  filter(year >= 2000 & Pos == "RB") %>%
+  ggplot(mapping=aes(x=Pick, y=CarAV)) +
+    geom_point()  +
+    labs(x = "Pick", y = "Career Value", title = "RB Average Value") + 
     geom_smooth(method='loess')
@@
 wr_av_pick <- master_df %>%
-  filter(year >= 2000) %>%
-  filter(Pos == "WR") %>%
+  filter(year >= 2000 & Pos == "WR") %>%
   ggplot(mapping=aes(x=Pick, y=CarAV)) +
     geom_point()  +
-    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for WRs") + 
+    labs(x = "Pick", y = "Career Value", title = "WR Average Value") + 
     geom_smooth(method='loess')
 
-grid.arrange(qb_av_pick, rb_av_pick, wr_av_pick, ncol=3)
+grid.arrange(qb_av_pick, rb_av_pick, wr_av_pick, ncol=3, top = "Average Value Over Picks in the 2000s")
 
@@
 
-grid.arrange(qb_violin, rb_violin, wr_violin, ncol=3)
+grid.arrange(qb_violin, rb_violin, wr_violin, ncol=3, top = "Average Value over Round")
 ```
@@
 Now, right off the bat, we hope that the model will predict that Joe Burrow has a higher likelihood of winning the MVP.
+Not only because he's a stud, but because he's fresh off the Heisman, and the first overall pick QB that is going to a team in desperate need of good QB play.
+Vegas has him as the clear favorite to win, and the season is set up for him to take the title, so if our model is worth anything
 

### 2020-05-11 draft_analysis.Rmd
Some reformatting and updated prose

@@
 library(gridExtra)
+library(caret)
+library(knitr)
 setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Draft Analytics/")
@@
 We will filter out any selections after round 7, as since 1994, the NFL draft has only consisted of 7 rounds.
+
+The result will be stored in the dataframe *master_df* with entities for each player.
+
+Note: We will also do some process in this step as well, but we will go into more detail on that when we repeat the processing steps in another dataframe later in the walkthrough.
 
@@
 
-We will also scrape the data from the 2020 draft for use in our machine learning problem and save it to our df variable.
+We will also scrape the data from the 2020 draft in the same way for use in our machine learning problem and save it to our *df_2020* variable.
 
@@
 
-Load the NFL draft history data, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft. Remove all of the missing data entries, and convert the necessary columns to integers. 
-```{r cars}
-head(master_df)
+Load the NFL draft history data from *master_df*, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft.
+Remove all of the missing data entries, and convert the necessary columns to integers.
+
+```{r process_master}
 
@@
 master_df$Pick <- as.integer(master_df$Pick)
+
+kable(head(master_df))
 ```
+
+Our *master_df* dataframe now has entities for each player with the following variables:
+
+    Rnd : Round in which the player was selected
+
+    Pick: Overall selection in draft
+
+    Tm  : Team
+
+    Player: Player name
+
+    Pos : Position
+
+    Age
+
+    To: Last year in the league
+
+    AP1: First team All-Pro selections
+
+    PB: Pro Bowl selections
+
+    St: Number of years as primary starter for his team at his position
+
+    CarAv: Weighted Career Approximate Value
+
+                  = 100% of AV of his best season + 95% of AV of his next best season + 90% of AV of his third-best season + ...
+
+    DrAV: Average value accumulated for the team that drafted this player
+
+    G : Games Played
+
+    year: Year drafted
 
@@
 First, we will get rid of columns we don't care about, then remove blank rows, add year drafted to each observation, and convert the necessary columns to integers, all like we did for the historical draft dataset.
-We will also filter the data to only include quarterbacks, wide receivers, and running backs.
+We will also filter this dataframe to only include quarterbacks, wide receivers, and running backs.
 
@@
 df_2020$Pos <- as.factor(df_2020$Pos)
+
+kable(head(df_2020))
 ```
+
+So, now our *df_2020* dataframe contains only the offensive skill position players that were drafted in 2020 as entities.
+The variables of this dataframe are **Rnd**, **Pick**, **Tm**, **Player**, **Pos**, **Age**, and **Year** like we have in our *master_df* variable.
 
@@
 Then for each period, we will take average career value of each position grouped by each round that the player was selected.
-We will make a new pos_df dataframe with these groupings.
+We will make a new *pos_df* dataframe with these groupings.
 
@@
 
-head(pos_df)
+kable(head(pos_df))
 ```
 
-Let's also add a dataframe that contains a variable that is TRUE when the player that was drafted was Offensive Rookie of the Year, and FALSE otherwise.
-We will also filter out all of the defensive players in this new dataframe.
+Let's also add a dataframe *off_df* that is the same as *master_df*, but contains a variable **roty** that is Yes when the player that was drafted from 1970-2016 was Offensive Rookie of the Year, and No otherwise.
+We will also filter the dataframe to only contain offensive skill position players, like we did for the 2020 draft data.
 
@@
 off_df$roty <- as.factor(off_df$roty)
+
+kable(head(off_df))
 ```
+
 # Part 3: Exploratory Data Analysis
 
-## Trends in Average value
+## Trends in Average Value
+
 Has average value for players selected in different rounds gone up over time?
@@
 
-## Analyzing the Quarterback position
+## Analyzing the Quarterback Position
 
@@
 
-## Analyzing the WR and RB positions
+## Analyzing the WR and RB
 
@@
 Not only because he's a stud, but because he's fresh off the Heisman, and the first overall pick QB that is going to a team in desperate need of good QB play.
-Vegas has him as the clear favorite to win, and the season is set up for him to take the title, so if our model is worth anything
+Vegas has him as the clear favorite to win, and the season is set up for him to take the title, so if our model is worth anything, it should at the very least have Joe Burrow as one of its top predictions for the award.
 
 First, lets control the resampling of our data.
-In this case, we’re going to cross-validate the data 3 times, therefore training it 3 times on different portions of the data before settling on the best tuning parameters (for gbm it is trees, shrinkage, and interaction depth).
+In this case, we’re going to cross-validate the data 3 times, therefore training the model 3 times on different portions of the data before settling on the best tuning parameters.
+For gbm, the tuning parameters are trees, shrinkage, and interaction depth.
 
 ```{r ml_control}
-
-library(caret)
-
-outcome_name <- 'roty'
-predictor_names <- c('Tm', 'Pos', 'Pick')
 
@@
 # Creating our Model
+
 Now to actually train the model...
 
+First, we will set the predictor variables to **Tm**, **Pos**, and **Pick**, and the sole outcome variable as **roty**.
+Then we will use the *train()* method from the caret package to train the model using the *off_df* variable that contains all our historical draft data on all the offensive skill position players.
+
 ```{r train_model}
+
+outcome_name <- 'roty'
+predictor_names <- c('Tm', 'Pos', 'Pick')
 
@@
 
+We will print the summary of the model below, which will display the relative influences of each of the predictor variables.
+
 ```{r sum_model}
@@
 
-The summaries tell us how much influence each variable had on the model.
+The summaries tell us how much relative influence each variable had on the model.
 It was interesting to see that position had such absolutely zero influence in predicting the model.
@@
 What is even more interesting is that the team the player was drafted to seemed to influence the model more than the pick number itself!
-Before I had trained the model, I expected pick, position, and then team to have relative influence in that order!
+Before I had trained the model, I expected pick, position, and then team to have relative influence in that order.
 
@@
 predictions <- predict(object=model, df_2020[,predictor_names], type='prob')
-head(predictions)
+kable(head(predictions))
 ```
@@
 
-Now let's get the top 7 players that the model is most certain will be the Offensive Rookie of the Year
+Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year
 
@@
 
-top10_roty
+kable(top10_roty)
 ```
@@
 We also see that Cam Akers, Tua, Henry Ruggs III, Jerry Jeudy, CeDee Lamb, and Justin Jefferson also appear in the top 10.
-All of these players currently have the highest odds to win the the Offensive MVP according to OddsShark.
+All of these players are among those that have the highest odds to win the the Offensive MVP according to OddsShark.
+
+Again, it's important to highlight that we trained this model JUST based on the pick number, team, and position for each player, whereas the Vegas odds have most certainly taken into account the player's performance in college as well.

### 2020-05-11 draft_analysis.Rmd
Add testing for ML model

@@
 
-```{r scraping}
+```{r scrape_master}
 
@@
 
-```{r scrape2020}
+```{r scrape_2020}
 
@@
 
-Next we will scrape the AP Offensive Rookie of the Year Winners from the same website.
+We will also scrape the 2017-2019 NFL draft data into a seperate dataframe *test_df*, which will be used to test our machine learning model that we will develop later in the tutorial.
+
+```{r scrape_test}
+test_df <- data.frame()
+for (year in 2017:2019){
+
+  draft_url <- paste("https://www.pro-football-reference.com/years/", year, "/draft.htm", sep="")
+  draft_html <- read_html(draft_url)
+
+  df <- draft_html %>%
+    html_node("table") %>%
+    html_table()
+
+  names(df) <- as.character(unlist(df[1,]))
+  df <- df[-1,]
+
+  # get rid of the columns we dont care about
+  df <- df[,1:6]
+
+  # add year drafted to each observation
+  df <- df %>%
+    mutate(year = year)
+
+  test_df <- rbind(test_df, df)
+}
+
+```
+
+Next we will scrape the AP Offensive Rookie of the Year Winners from the same website, again for our machine learning problem.
 
@@
 The variables of this dataframe are **Rnd**, **Pick**, **Tm**, **Player**, **Pos**, **Age**, and **Year** like we have in our *master_df* variable.
+
+We will process the *test_df* dataframe in the same way.
+
+```{r process_test}
+
+# filter out certain rounds
+test_df <- test_df %>%
+  filter(Rnd < 8)
+
+# get rid of blank rows
+test_df <-test_df[!(test_df$Player=="Player"),]
+
+# filter so dataframe only contains offensive skill position players.
+test_df <- test_df %>%
+  filter(Pos == "QB" | Pos == "WR" | Pos == "RB")
+
+test_df[is.na(test_df)] <- 22
+
+test_df$Pick <- as.integer(test_df$Pick)
+test_df$Age <- as.integer(test_df$Age)
+test_df$Tm <- as.factor(test_df$Tm)
+test_df$Pos <- as.factor(test_df$Pos)
+
+kable(head(test_df))
+```
 
@@
 
-# Predicting the 2020 Rookie Offensive MVP
+Let's test our model against the 2017, 2018, and 2019 datasets.
+To do this, we will use the *predict()* function in the caret package and feed it the test data.
+From there,  *predict()* will calculate probabilities that the model is sure about it choice that each observation from *test_df* will fall under the "Yes" and "No" outcome categories.
+So the "Yes" variable corresponds to the probability the model is sure about its choice that the player will become Rookie MVP, and "No" = probability the model is sure the player will not be the Rookie MVP.
 
-Now let's use the model to predict the probabilities that each player from the 2020 draft will become Offensive Rookie of the Year.
-This code will give us the probability that the model is sure about its choice that the observation will fall under the "No" and "Yes" category.
+```{r eval_test}
 
-```{r eval}
-
-predictions <- predict(object=model, df_2020[,predictor_names], type='prob')
+predictions <- predict(object=model, test_df[,predictor_names], type='prob')
 kable(head(predictions))
@@
 
+```{r join_probabilities_test}
+
+test_df <- merge(test_df, predictions, by = "row.names", all.x = FALSE, all.y = FALSE)
+test_df$No.x <- NULL
+
+kable(head(test_df))
+```
+
+Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year for 2017, 2018, and 2019.
+
+```{r top10_roty_test}
+
+top10_roty_2017 <- test_df %>%
+  filter(year == 2017) %>%
+  top_n(10, Yes)
+
+kable(top10_roty_2017)
+
+top10_roty_2018 <- test_df %>%
+  filter(year == 2018) %>%
+  top_n(10, Yes)
+
+kable(top10_roty_2018)
+
+top10_roty_2019 <- test_df %>%
+  filter(year == 2019) %>%
+  top_n(10, Yes)
+
+kable(top10_roty_2019)
+```
+
+The actual winners for Rookie Offensive MVP for 2017-2019 were Alvin Kamara, Saquon Barkley, and Kyler Murray (in that order).
+The model fails to list Kamara as a top 10 prediction for 2017, but lists Barkley as its fourth best prediction for 2018, and Murray as its top prediction for 2019!
+
+# Predicting the 2020 Rookie Offensive MVP
+
+Now let's use the model to predict the probabilities that each player from the 2020 draft will become Offensive Rookie of the Year.
+Like we did for the testing portion, this code will give us the probability that the model is sure about its choice that the observation will fall under the "No" and "Yes" category.
+
+```{r eval_2020}
+
+predictions <- predict(object=model, df_2020[,predictor_names], type='prob')
+```
+
+Now let's join our prediction probabilities to the actual data like we did before. s
+
 ```{r join_probabilities}
@@
 df_2020$No.x <- NULL
+
+kable(head(df_2020))
 ```

### 2020-05-12 draft_analysis.Rmd
Reformatted headings
Delete unnecessary columns

@@
 
-# Creating our Model
+## Creating our Model
 
@@
 
-# Testing against 2017-2019 data
+## Testing against 2017-2019 data
 
@@
 test_df$No.x <- NULL
+test_df$Row.names <- NULL
 
@@
 
-# Predicting the 2020 Rookie Offensive MVP
+## Predicting the 2020 Rookie Offensive MVP
 
@@
 df_2020$No.x <- NULL
+df_2020$Row.names <- NULL
 

### 2020-05-12 draft_analysis.Rmd
Add lm line to all the graphs (in addition to the loess lines)

@@
     labs(x = "Pick", y = "Career Value", title = "Average Value over picks for QBs Drafted in the 2000s") + 
-    geom_smooth(method='loess')
+    geom_smooth(method='loess') + 
+    geom_smooth(method=lm, color="green")
 
@@
     labs(x = "Pick", y = "Career Value", title = "QB Average Value") + 
-    geom_smooth(method='loess')
+    geom_smooth(method='loess') + 
+    geom_smooth(method=lm, color="green")
 
@@
     labs(x = "Pick", y = "Career Value", title = "RB Average Value") + 
-    geom_smooth(method='loess')
+    geom_smooth(method='loess') + 
+    geom_smooth(method=lm, color="green")
 
@@
     labs(x = "Pick", y = "Career Value", title = "WR Average Value") + 
-    geom_smooth(method='loess')
+    geom_smooth(method='loess') + 
+    geom_smooth(method=lm, color="green")
 

### 2020-05-12 draft_analysis.Rmd
Reformatted titles for some graphs

@@
 
-grid.arrange(qb_bar, rb_bar, wr_bar, ncol=3, top = "Average Value in Different Rounds over the 2000s")
+grid.arrange(qb_bar, rb_bar, wr_bar, ncol=3, top = "Average Value in Different Rounds in the 2000s")
 ```
@@
 
+qb_violin <- master_df %>%
+  filter(year >= 2000) %>%
+  filter(Pos == "QB") %>%
+  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
+    geom_violin()  +
+    labs(x = "Rnd", y = "Career Value", title = "QB Average Value") + 
+    geom_smooth(method=lm)
+
 rb_violin <- master_df %>%
@@
     geom_violin()  +
-    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for RBs") + 
+    labs(x = "Rnd", y = "Career Value", title = "RB Average Value") + 
     geom_smooth(method=lm)
@@
     geom_violin()  +
-    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for WRs") + 
+    labs(x = "Rnd", y = "Career Value", title = "WR Average Value") + 
     geom_smooth(method=lm)
 
-grid.arrange(qb_violin, rb_violin, wr_violin, ncol=3, top = "Average Value over Round")
+grid.arrange(qb_violin, rb_violin, wr_violin, ncol=3, top = "Average Value over Round in the 2000s")
 ```

### 2020-05-12 draft_analysis.Rmd
Fix working directory

@@
 library(knitr)
-setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Draft Analytics/")
+setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 ```

### 2020-05-12 draft_analysis.Rmd
Remove working dir stuff (its not necessary!)

@@
 library(knitr)
-setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
 ```

### 2020-05-12 draft_analysis.Rmd
Updated prose to add more details

@@
 
-Next we will scrape the AP Offensive Rookie of the Year Winners from the same website, again for our machine learning problem.
+Next we will scrape the AP Offensive Rookie of the Year Winners from the same website, again for our machine learning problem, and save into the *roty_df* dataframe.
 
@@
 Load the NFL draft history data from *master_df*, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft.
-Remove all of the missing data entries, and convert the necessary columns to integers.
+We will only keep the observations that do not have any missing data entries.
+Also, we will convert the necessary columns to integers and factors (rather than chars).
+This way, we can use the columns values to group and to produce regression lines in our data visualization steps later.
 
@@
 Now let's process our 2020 Draft data.
-First, we will get rid of columns we don't care about, then remove blank rows, add year drafted to each observation, and convert the necessary columns to integers, all like we did for the historical draft dataset.
+First, we will get rid of columns we don't care about, then remove blank rows, add year drafted to each observation, and convert the necessary columns to integers and factors, all like we did for the historical draft dataset.
 We will also filter this dataframe to only include quarterbacks, wide receivers, and running backs.
@@
 
-We will process the *test_df* dataframe in the same way.
+We will process the *test_df* dataframe in the same way as the *df_2020* dataframe.
 
@@
 
+Our first visualization will answer the following questions for us:
 Has average value for players selected in different rounds gone up over time?
 Have GMs gotten better at drafting in the first round, second round, etc...?
+
+To do this, we will plot the average value for all the players for each year since 1970, grouped by round.
 
@@
 
+So, lets create a bar graph. For each time period, we will graph the average value of a quarterback selected in each round.
+
 ```{r qb}
@@
 Now let's focus our attention on the past 20 years and graph some visualizations of that data.
+
+Our first graph for the 2000s will be the average value of a quarterback on the y-axis, and the pick number in which they were selected on the x-axis.
+
 ```{r qb2}
@@
 This would suggest that seelcting a QB at picks 100-150 would, statistically speaking, provide less bang for your buck, as you could likely get a player of the same quality with some of your later draft capital.
+
+Our next plot will we a violin plot of the average value for quarterbakcs drafted in the 2000s in different rounds.
+This will give us a better idea of the distribution of the average value for the quarterback position.
 
@@
 
-Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft.
+Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft. We will use *grid.arrange* to show the graphs side by side in three columns.
 
@@
 From there,  *predict()* will calculate probabilities that the model is sure about it choice that each observation from *test_df* will fall under the "Yes" and "No" outcome categories.
-So the "Yes" variable corresponds to the probability the model is sure about its choice that the player will become Rookie MVP, and "No" = probability the model is sure the player will not be the Rookie MVP.
+So the **Yes** variable corresponds to the probability the model is sure about its choice that the player will become Rookie MVP, and **No** = probability the model is sure the player will not be the Rookie MVP.
 
@@
 Now let's join our prediction probabilities to the actual data, so we can look at the prediction probabilities for each player.
+We will use *merge*, which concatentaes dataframes horizontally.
+Setting the "by" parameter to "row.names" matches the dataframes by row index.
 
@@
 
-Now let's join our prediction probabilities to the actual data like we did before. s
+Now let's join our prediction probabilities to the actual data like we did before.
 
@@
 
-Joe Burrow is our model's top prediction to win the MVP!!!
-We also see that Cam Akers, Tua, Henry Ruggs III, Jerry Jeudy, CeDee Lamb, and Justin Jefferson also appear in the top 10.
+**Joe Burrow is our model's top prediction to win the MVP!!!** We also see that Cam Akers, Tua, Henry Ruggs III, Jerry Jeudy, CeDee Lamb, and Justin Jefferson also appear in the top 10.
 All of these players are among those that have the highest odds to win the the Offensive MVP according to OddsShark.

### 2020-05-13 draft_analysis.Rmd
Updated tutorial and added some supporting links

@@
 
-The result will be stored in the dataframe *master_df* with entities for each player.
+To scrape the data, we will use the *read_html()*, *html_node()*, and *html_table()* methods in the **rvest** package.
+More information on **rvest** is available here: https://blog.rstudio.com/2014/11/24/rvest-easy-web-scraping-with-r/
 
-Note: We will also do some process in this step as well, but we will go into more detail on that when we repeat the processing steps in another dataframe later in the walkthrough.
+First, we store the html file from our draft data website into the *draft_html* variable.
+From there, we search for the first occurence of the "table" element using *html_node()*, and then we parse the HTML data results into a dataframe using the *html_table()*.
+After parsing, we see that the column names are in the first row of the data, so we set the column names to be the first row of data with *names(df) <- as.character(unlist(df[1,]))*.
+Then we delete the first row of data (that still contains the column names) so all of the rows just contain player data.
+
+Since each years draft data is on a different website, we loop through all of the website, and after some more processing (see note), we use *rbind()* to add each dataframe vertically to our *master_df* dataframe.
+
+The result will be all of the 1970-2016 draft data stored in the dataframe *master_df* with entities for each player.
+
+*Note: We will also do some process in this step as well, but we will go into more detail on that when we repeat the processing steps in another dataframe later in the walkthrough.
 
@@
 Load the NFL draft history data from *master_df*, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft.
-We will only keep the observations that do not have any missing data entries.
-Also, we will convert the necessary columns to integers and factors (rather than chars).
+The first line in this chunk will ensure that we will only keep the observations that do not have any missing data entries.
+Then, we will convert the necessary columns to integers and factors (rather than chars).
 This way, we can use the columns values to group and to produce regression lines in our data visualization steps later.
@@
 Now let's process our 2020 Draft data.
-First, we will get rid of columns we don't care about, then remove blank rows, add year drafted to each observation, and convert the necessary columns to integers and factors, all like we did for the historical draft dataset.
-We will also filter this dataframe to only include quarterbacks, wide receivers, and running backs.
+We will be using the *filter()* and *mutate()* methods found in the **dplyr** package in the **tidyverse** package.
+More info on **dplyr** can be found here: https://dplyr.tidyverse.org/
+
+First, we will get rid of columns we don't care about. 
+Next, we only keep rounds 1-7 using *filter()* with the conditional that **Rnd** variable must be less than 8. 
+There seem to be seperator rows between rounds that contain the column names again (probably for readability on the website), so we get rid of them by choosing to keep only rows where the **Player** variable does not equal "Player".
+We then add year drafted to each observation using *mutate()*, which allows us to add variables to an existing dataframe, and again use *filter()* to only include quarterbacks, wide receivers, and running backs.
+We noticed that some of the age data is missing for the players, so we fill in all of the missing age values with 22, since that is the most common age for a player to enter the pros.
+We convert the necessary columns to integers and factors, all like we did for the historical draft dataset.
 
@@
 
-# get rid of blank rows
+# get rid of seperator rows
 df_2020 <-df_2020[!(df_2020$Player=="Player"),]
@@
 First, we want to break up our data into 4 periods.
-Then for each period, we will take average career value of each position grouped by each round that the player was selected.
+We accomplish this with the *cut()* method on our **year** variable with parameter 'breaks' = 4.
+This will break up all of the data into 4 periods based on year, more information on cut is available here: https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cut
+
+We will use *mutate()* again to add a variable **period** for which period this entry is in.
+Then for each period, we will take average career value of each position using *summarise()*, grouped by each round, period and position that the player was selected using *group_by()*.
+Both of these methods are also available in the **dplyr** package.
+
 We will make a new *pos_df* dataframe with these groupings.
@@
 Let's also add a dataframe *off_df* that is the same as *master_df*, but contains a variable **roty** that is Yes when the player that was drafted from 1970-2016 was Offensive Rookie of the Year, and No otherwise.
+Again, since we are adding a new variable, we use *mutate()*, with a call to *ifelse()*.
+The condition in the ifelse will check if the entry's player name exists in the **Player** column of our rookie of the year table, if it exists it returns "Yes", and "No" otherwise.
+Lastly, we convert this new variable into a factor w/ 2 levels.
+
 We will also filter the dataframe to only contain offensive skill position players, like we did for the 2020 draft data.
@@
 To do this, we will plot the average value for all the players for each year since 1970, grouped by round.
+We will create our plots using **ggplot2** which is available in the **tidyverse** package, more info can be found here: https://ggplot2.tidyverse.org/
 
@@
 As we would expect the trend is that there is a decline in value until about pick 100, where the career value seems to be about the same from about pick 100-250.
-This would suggest that seelcting a QB at picks 100-150 would, statistically speaking, provide less bang for your buck, as you could likely get a player of the same quality with some of your later draft capital.
+So the data suggests that selecting a QB at picks 100-150 would provide less bang for your buck, as you could likely get a player of the same quality with some of your later draft capital.
 
@@
 
+All of our machine learning goodness uses methods from the **caret** package, more info on this package can be found here: https://cran.r-project.org/web/packages/caret/vignettes/caret.html
+
 ## Predicting the Rookie Offensive MVP
@@
 First, we will set the predictor variables to **Tm**, **Pos**, and **Pick**, and the sole outcome variable as **roty**.
-Then we will use the *train()* method from the caret package to train the model using the *off_df* variable that contains all our historical draft data on all the offensive skill position players.
+Then we will use the *train()* method from the **caret** package to train the model using the *off_df* variable that contains all our historical draft data on all the offensive skill position players.
+We tell *train()* what predictor and outcome variables to use, as well as to use the gbm method with the resampling control that we set previously.
 
@@
 Let's test our model against the 2017, 2018, and 2019 datasets.
-To do this, we will use the *predict()* function in the caret package and feed it the test data.
-From there,  *predict()* will calculate probabilities that the model is sure about it choice that each observation from *test_df* will fall under the "Yes" and "No" outcome categories.
+To do this, we will use the *predict()* function in the **caret** package and feed it the test data.
+Since we set parameter 'type' to "prob",  *predict()* will calculate probabilities that the model is sure about it choice that each observation from *test_df* will fall under the "Yes" and "No" outcome categories.
 So the **Yes** variable corresponds to the probability the model is sure about its choice that the player will become Rookie MVP, and **No** = probability the model is sure the player will not be the Rookie MVP.
@@
 
-Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year for 2017, 2018, and 2019.
+Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year for 2017, 2018, and 2019 respectively.
 
@@
 
-Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year
+Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year in 2020.
 

### 2020-05-15 draft_analysis.Rmd
Updated prose

@@
 
-Load the NFL draft history data from *master_df*, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft.
+First, we will tidy the *master_df* dataframe.
 The first line in this chunk will ensure that we will only keep the observations that do not have any missing data entries.
@@
 
-Now let's process our 2020 Draft data.
+Now let's process/tidy our 2020 Draft data.
 We will be using the *filter()* and *mutate()* methods found in the **dplyr** package in the **tidyverse** package.
@@
 We will use *mutate()* again to add a variable **period** for which period this entry is in.
-Then for each period, we will take average career value of each position using *summarise()*, grouped by each round, period and position that the player was selected using *group_by()*.
+Then for each grouping of period, round, and period using *group_by()*, we will create an average career value variable  **Avg_AV** of each grouping using *summarise()*.
 Both of these methods are also available in the **dplyr** package.
@@
 
+Here, we map the x-variable to **year** and the y-variable to **Avg_AV**, and using *geom_point()* tells **ggplot2** to create a scatter plot of the data.
+We them use *geom_smooth(method=lm)* to create a regression line, where lm stands for linear model.
+More info can be found here: https://ggplot2.tidyverse.org/reference/geom_smooth.html
+
+We group by round to create 7 different regression lines in our plot.
+
 ```{r avg_val_rd1}
@@
 
+To create a bar graph, we use *geom_bar*, where stat="identity" allows us to have the height of the bars represent values in the data, which are mapped to **Avg_AV** in this case.
+The position="dodge" puts the bars for each round side by side.
+
+More info on *geom_bar()*: https://ggplot2.tidyverse.org/reference/geom_bar.html
+
 ```{r qb}
@@
 
-Now let's focus our attention on the past 20 years and graph some visualizations of that data.
+Now let's focus our attention on QBs of the past 20 years and graph some visualizations of that data. To do this, we will use *filter()* on **Pos** and **year**.
 
 Our first graph for the 2000s will be the average value of a quarterback on the y-axis, and the pick number in which they were selected on the x-axis.
+
+In this case, in addition to the linear model regression line in green, we will also add a locally estimated scatterplot smoothing (LOESS) regression line using *geom_smooth(method='loess')*.
 
@@
 
+We use *geom_violion()* to create the violin plots, more info can be found here: https://ggplot2.tidyverse.org/reference/geom_violin.html
+
 ```{r qb3}
@@
     geom_violin()  +
-    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for QBs Drafted in the 2000s") + 
-    geom_smooth(method=lm)
-
+    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for QBs Drafted in the 2000s")
 qb_violin
@@
 
-Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft. We will use *grid.arrange* to show the graphs side by side in three columns.
+Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft.
+We will use *grid.arrange* to show the graphs for WRs, RBs, and QBs side by side in three columns.
 
@@
 
-Not as drastic of a difference in rounds 1 and 2 in the WR group. The RB average value in the 2000s seems to resemble the QB position.
+Insights:
+Not as drastic of a difference in rounds 1 and 2 in the WR group. 
+The RB average value in the 2000s seems to resemble the QB position.
 
@@
 
+Insights:
 Much less steep of a drop off between round 1 to 2 in the WR group.
-
 Again we see a trend that by pick 100 the value seems to hit a plateue.
@@
     geom_violin()  +
-    labs(x = "Rnd", y = "Career Value", title = "QB Average Value") + 
-    geom_smooth(method=lm)
+    labs(x = "Rnd", y = "Career Value", title = "QB Average Value")
 
@@
     geom_violin()  +
-    labs(x = "Rnd", y = "Career Value", title = "RB Average Value") + 
-    geom_smooth(method=lm)
+    labs(x = "Rnd", y = "Career Value", title = "RB Average Value")
 
@@
     geom_violin()  +
-    labs(x = "Rnd", y = "Career Value", title = "WR Average Value") + 
-    geom_smooth(method=lm)
+    labs(x = "Rnd", y = "Career Value", title = "WR Average Value")
 
@@
 
-RBs in Round 1 always produce.
+Insights:
+RBs in Round 1 always produce, and RBs selected in later rounds have a much larger spread than quarterbacks, suggesting it is easier to get a running back of decent value in later rounds than it is to get a QB.
 We again see that round 2 is a value round for WRs as it appears to resemble the type of value you get out of round 1.
+Rounds 1-4 have generally similar range in the spread of values, so there seems to be great talent available outside of rounds 1 and 2 for wide receivers.
 
@@
 First, lets control the resampling of our data.
-In this case, we’re going to cross-validate the data 3 times, therefore training the model 3 times on different portions of the data before settling on the best tuning parameters.
-For gbm, the tuning parameters are trees, shrinkage, and interaction depth.
+In this case, we’re going to cross-validate the data 3 times, so we will train the model 3 times on different portions of the data before deciding on the best tuning parameters.
+For the gbm method that we will be using, the tuning parameters are trees, shrinkage, and interaction depth.
 
@@
 objControl <- trainControl(method='cv', number=3, returnResamp='none', summaryFunction = twoClassSummary, classProbs = TRUE)
-
 ```
@@
 Then we will use the *train()* method from the **caret** package to train the model using the *off_df* variable that contains all our historical draft data on all the offensive skill position players.
-We tell *train()* what predictor and outcome variables to use, as well as to use the gbm method with the resampling control that we set previously.
+We tell *train()* what predictor and outcome variables to use, as well as to use the gbm (gradient boosting algorithm) with the resampling control that we set previously.
 
@@
 
-Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year for 2017, 2018, and 2019 respectively.
+So we see that the model is 85% percent sure that Mitchell Trubisky will not win Rookie Offensive MVP, and ~15% sure that Trubisky will win the award.
+
+Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year for 2017, 2018, and 2019 respectively, using the *top_n()* function in the **dplyr** package (https://dplyr.tidyverse.org/reference/top_n.html)
 
@@
 
-**Joe Burrow is our model's top prediction to win the MVP!!!** We also see that Cam Akers, Tua, Henry Ruggs III, Jerry Jeudy, CeDee Lamb, and Justin Jefferson also appear in the top 10.
+**Joe Burrow is our model's top prediction to win the MVP!!!** The model is 21% sure that Burrow will win the Rookie Offensive MVP, and 79% sure that he will not.
+
+We also see that Cam Akers, Tua, Henry Ruggs III, Jerry Jeudy, CeDee Lamb, and Justin Jefferson also appear in the top 10.
 All of these players are among those that have the highest odds to win the the Offensive MVP according to OddsShark.
